
```{r}
library(tidyverse)
library(fpp3)
library(GGally)
```

```{r}
PBS %>%
  filter(ATC2 == "A10") %>%
  select(Month, Concession, Type, Cost) %>%
  summarise(TotalC = sum(Cost)) %>%
  mutate(Cost = TotalC / 1e6) -> a10
```

Suppose you have annual observations for the last few years:

```{r}
prison <- readr::read_csv("https://OTexts.com/fpp3/extrafiles/prison_population.csv")
```
We can read it into R, and create a tsibble object, by simply identifying which column contains the time index, and which columns are keys. The remaining columns are values — there can be many value columns, although in this case there is only one (Count). The original csv file stored the dates as individual days, although the data is actually quarterly, so we need to convert the Date variable to quarters.

```{r}
prison <- prison %>%
  mutate(Quarter = yearquarter(Date)) %>%
  select(-Date) %>%
  as_tsibble(key = c(State, Gender, Legal, Indigenous),
             index = Quarter)
```

For time series data, the obvious graph to start with is a time plot. That is, the observations are plotted against the time of observation, with consecutive observations joined by straight lines. Figure 2.1 shows the weekly economy passenger load on Ansett airlines between Australia’s two largest cities.

```{r}
melsyd_economy <- ansett %>%
  filter(Airports == "MEL-SYD", Class == "Economy") %>%
  mutate(Passengers = Passengers/1000)

autoplot(melsyd_economy, Passengers) +
  labs(title = "Ansett airlines economy class",
       subtitle = "Melbourne-Sydney",
       y = "Passengers ('000)")
```
```{r}
autoplot(a10, Cost) +
  labs(y = "$ (millions)",
       title = "Australian antidiabetic drug sales")
```
A seasonal plot is similar to a time plot except that the data are plotted against the individual “seasons” in which the data were observed. An example is given in Figure 2.4 showing the antidiabetic drug sales.

```{r}
a10 %>%
  gg_season(Cost, labels = "both") +
  labs(y = "$ (millions)",
       title = "Seasonal plot: Antidiabetic drug sales")
```
These are exactly the same data as were shown earlier, but now the data from each season are overlapped. A seasonal plot allows the underlying seasonal pattern to be seen more clearly, and is especially useful in identifying years in which the pattern changes.

In this case, it is clear that there is a large jump in sales in January each year. Actually, these are probably sales in late December as customers stockpile before the end of the calendar year, but the sales are not registered with the government until a week or two later. The graph also shows that there was an unusually small number of sales in March 2008 (most other years show an increase between February and March). The small number of sales in June 2008 is probably due to incomplete counting of sales at the time the data were collected.

Multiple seasonal periods

Where the data has more than one seasonal pattern, the period argument can be used to select which seasonal plot is required. The vic_elec data contains half-hourly electricity demand for the state of Victoria, Australia. We can plot the daily pattern, weekly pattern or yearly pattern by specifying the period argument as shown in Figures 2.5–2.7.

In the first plot, the three days with 25 hours are when daylight saving ended in each year and so these days contained an extra hour. There were also three days with only 23 hours each (when daylight saving started) but these are hidden beneath all the other lines on the plot.

Dia

```{r}
vic_elec %>% gg_season(Demand, period = "day") +
  theme(legend.position = "none") +
  labs(y="MWh", title="Electricity demand: Victoria")
```

Semana

```{r}
vic_elec %>% gg_season(Demand, period = "week") +
  theme(legend.position = "none") +
  labs(y="MWh", title="Electricity demand: Victoria")
```

Ano

```{r}
vic_elec %>% gg_season(Demand, period = "year") +
  labs(y="MWh", title="Electricity demand: Victoria")
```

An alternative plot that emphasises the seasonal patterns is where the data for each season are collected together in separate mini time plots.

```{r}

a10 %>%
  gg_subseries(Cost) +
  labs(
    y = "$ (millions)",
    title = "Australian antidiabetic drug sales"
  )
```
Australian quarterly vacation data provides an interesting example of how these plots can reveal information. First we need to extract the relevant data from the tourism tsibble. All the usual tidyverse wrangling verbs apply. To get the total visitor nights spent on Holiday by State for each quarter (i.e., ignoring Regions) we can use the following code. Note that we do not have to explicitly group by the time index as this is required in a tsibble.

```{r}
holidays <- tourism %>%
  filter(Purpose == "Holiday") %>%
  group_by(State) %>%
  summarise(Trips = sum(Trips))
```

Time plots of each series show that there is strong seasonality for most states, but that the seasonal peaks do not coincide.
```{r}
autoplot(holidays, Trips) +
  labs(y = "Overnight trips ('000)",
       title = "Australian domestic holidays")
```
To see the timing of the seasonal peaks in each state, we can use a season plot. Figure 2.10 makes it clear that the southern states of Australia (Tasmania, Victoria and South Australia) have strongest tourism in Q1 (their summer), while the northern states (Queensland and the Northern Territory) have the strongest tourism in Q3 (their dry season).

```{r}
gg_season(holidays, Trips) +
  labs(y = "Overnight trips ('000)",
       title = "Australian domestic holidays")
```
```{r}
holidays %>%
  gg_subseries(Trips) +
  labs(y = "Overnight trips ('000)",
       title = "Australian domestic holidays")
```
The graphs discussed so far are useful for visualising individual time series. It is also useful to explore relationships between time series.

Figures 2.12 and 2.13 show two time series: half-hourly electricity demand (in Gigawatts) and temperature (in degrees Celsius), for 2014 in Victoria, Australia. The temperatures are for Melbourne, the largest city in Victoria, while the demand values are for the entire state.


```{r}
vic_elec %>%
  filter(year(Time) == 2014) %>%
  autoplot(Demand) +
  labs(y = "GW",
       title = "Half-hourly electricity demand: Victoria")
```
```{r}
vic_elec %>%
  filter(year(Time) == 2014) %>%
  autoplot(Temperature) +
  labs(
    y = "Degrees Celsius",
    title = "Half-hourly temperatures: Melbourne, Australia"
  )
```
```{r}
vic_elec %>%
  filter(year(Time) == 2014) %>%
  autoplot(Temperature) +
  labs(
    y = "Degrees Celsius",
    title = "Half-hourly temperatures: Melbourne, Australia"
  )
```

We can study the relationship between demand and temperature by plotting one series against the other.

```{r}
vic_elec %>%
  filter(year(Time) == 2014) %>%
  ggplot(aes(x = Temperature, y = Demand)) +
  geom_point() +
  labs(x = "Temperature (degrees Celsius)",
       y = "Electricity demand (GW)")
```
This scatterplot helps us to visualise the relationship between the variables. It is clear that high demand occurs when temperatures are high due to the effect of air-conditioning. But there is also a heating effect, where demand increases for very low temperatures.

When there are several potential predictor variables, it is useful to plot each variable against each other variable. Consider the eight time series shown in Figure 2.17, showing quarterly visitor numbers across states and territories of Australia.


```{r}
visitors <- tourism %>%
  group_by(State) %>%
  summarise(Trips = sum(Trips))
visitors %>%
  ggplot(aes(x = Quarter, y = Trips)) +
  geom_line() +
  facet_grid(vars(State), scales = "free_y") +
  labs(title = "Australian domestic tourism",
       y= "Overnight trips ('000)") +
  guides(size = F)
```

To see the relationships between these eight time series, we can plot each time series against the others. These plots can be arranged in a scatterplot matrix, as shown in Figure 2.18. (This plot requires the GGally package to be installed.)

```{r}
visitors %>%
  pivot_wider(values_from=Trips, names_from=State) %>%
  GGally::ggpairs(columns = 2:9)
```


displays scatterplots of quarterly Australian beer production (introduced in Figure 1.1), where the horizontal axis shows lagged values of the time series. Each graph shows  yt plotted against yt − k   for different values of  k .

```{r}
recent_production <- aus_production %>%
  filter(year(Quarter) >= 2000)
recent_production %>%
  gg_lag(Beer, geom = "point") +
  labs(x = "lag(Beer, k)")
```


The autocorrelation coefficients for the beer production data can be computed using the ACF() function.

```{r}
recent_production %>% ACF(Beer, lag_max = 9)
```
Just as correlation measures the extent of a linear relationship between two variables, autocorrelation measures the linear relationship between lagged values of a time series.
```{r}
recent_production %>%
  ACF(Beer) %>%
  autoplot() + labs(title="Australian beer production")
```
In this graph:

- r4  is higher than for the other lags. This is due to the seasonal pattern in the data: the peaks tend to be four quarters apart and the troughs tend to be four quarters apart.
- r2  is more negative than for the other lags because troughs tend to be two quarters behind peaks.
- The dashed blue lines indicate whether the correlations are significantly different from zero


When data have a trend, the autocorrelations for small lags tend to be large and positive because observations nearby in time are also nearby in value. So the ACF of a trended time series tends to have positive values that slowly decrease as the lags increase.

When data are seasonal, the autocorrelations will be larger for the seasonal lags (at multiples of the seasonal period) than for other lags.


When data are both trended and seasonal, you see a combination of these effects. The a10 data plotted in Figure  shows both trend and seasonality. Its ACF is shown in Figure 2.21. The slow decrease in the ACF as the lags increase is due to the trend, while the “scalloped” shape is due to the seasonality.

```{r}
a10 %>%
  ACF(Cost, lag_max = 48) %>%
  autoplot() +
  labs(title="Australian antidiabetic drug sales")
```

